---
  output: 
  md_document:
  variant: markdown_github
bibliography: paper.bib
---
  
# Summary

The `PID.POS` package is designed to aid with the identification of personal identifiability risks in data sets.
By applying existing natural language processing techniques, the package is able to identify proper nouns within a
data set for rapid manual review and redaction and hence reduce the operational burden of proper due diligence in data 
sharing.

# Statement of need  

The world is embedded in a data revolution.  Never before have we had the depth or breadth of data being captured 
and analysed than we do at present, and this is only set to increase.  Following a period of rapid  development 
in the spaces of machine learning and artificial intelligence, regulatory bodies are now playing catch up, with 
national and international acts coming into force to protect individual rights to privacy.  

One effect of increase legislation and visibility of privacy concerns has been a surge in data literacy,  with
data controllers becoming more aware of the risks associated with the data they hold.  Of all concerns, 'personal
identifiability' i.e. the ability to directly or indirectly identify an individual from a dataset, is paramount 
and, if breeched, can lead to reputation damage and fines. Where data is structured and  complete, a manual inspection 
can readily  identify variables which contain directly personally identifiable data (PID).  However, in the case of 
modern large data sets which may comprise millions of observations,  a manual inspection may miss PID if it is 
embedded within a passage of text, or is a rarity for the given variable.  

The `PID.POS` (Personal Identifiability Detection by Part Of Speech tagging) package is designed to aid 
with the identification of PID risks in data sets.   In comparison to existing packages which rely on a curated list
of common names and string-matching, `PID.POS` builds on the existing `udpipe` framework, extracting all examples of 
proper nouns and providing a mechanism for the review and redaction of PID risks.


# Comparison to existing R packages

The need to review data sets to identify risks is not new, and there are a number of packages which have been developed
to aid in this process.  The most notable of these are the `PII` package, which is designed to 
identify personally identifiable features via pattern matching. These approaches can be effective in 
identifying PID, but have a risk of missing edge cases e.g. relying on sentence case  to identify names. The 
approach taken in `PID.POS` conversely takes the approach of purposefully extracting all proper-nouns, and hence
increase the false positive rate, with the intention of supplying a simplified extract to aid  human interpretation 
rather than fully automate it.

# In practice

To assist with understanding the `PID.POS` package, we include a subset of the 'friends' 
data set from the `friends` package.

``` {r}
library(dplyr)

glimpse(the_one_in_massapequa)
```

``` {r}
report <- data_frame_report(the_one_in_massapequa)
report
```


