---
title: "Automatic Replacement Tools"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Automatic Replacement Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  rows.print = 6
)
library(pid.pos)

.tmp <- tempdir()
pid.pos:::set_model_folder(.tmp)

{
  model <- "english-ewt-ud-2.5-191206.udpipe"
  to <- file.path(.tmp, model)
  url <- file.path("https://raw.githubusercontent.com/jwijffels/udpipe.models.ud.2.5/master/inst/udpipe-ud-2.5-191206/", model)
  curl::curl_download(url, to, mode = "wb")
}
```

This package is designed to support the identification and redaction of personally identifiable data (PID) within textual datasets.

The `report_to_redaction_rules()` function generates a `replacement_rules` table from the output of `pid_pos()`. This table specifies candidate values for redaction and provides a `To` column in which users can define replacement values.

In many cases, users may wish to manually populate the `To` column (for example, replacing names with consistent pseudonyms). However, when a dataset contains a large volume of PID, manual specification may be impractical. In such cases, the `auto_replace()` utility can be used to automatically generate replacement values.

`auto_replace()` operates on a `replacement_rules` table and encodes the `To` column according to a user-defined replacement function. The package provides three built-in replacement strategies:

- `hashing_replacement.f()` — hashes values using a key and salt, producing deterministic and reproducible replacements.
- `random_replacement.f()` — generates random replacements from a defined character space.
- `all_random_replacement.f()` — generates random replacements while ensuring all outputs are unique.

---

# Basic Workflow

We begin by generating a PID report and converting it into redaction rules:

```{r, results="hide", message=FALSE, eval=F}
library(pid.pos)

report <- pid_pos(the_one_in_massapequa)
replacement_rules <- report_to_redaction_rules(report)
replacement_rules
```

```{r, echo=FALSE, message=FALSE}
report <- pid_pos(the_one_in_massapequa)
replacement_rules <- report_to_redaction_rules(report)

head(replacement_rules, 5)
```

To automatically populate the `To` column, we first initialise a replacement function.  
In this example, we use `random_replacement.f()` to generate five-character replacements drawn from uppercase letters:

```{r}
replacement.f <- random_replacement.f(
  replacement_size = 5,
  replacement_space = LETTERS
)
```

We then apply this function using `auto_replace()`:

```{r, results="hide"}
set.seed(101)
updated_replacement_rules <- auto_replace(
  replacement_rules,
  replacement.f
)

updated_replacement_rules
```

```{r, echo=FALSE}
head(updated_replacement_rules, 5)
```

The resulting `updated_replacement_rules`  can then be used alongside the original data in the `redact` function to adjust the data:

```{r}
redact(
  head(the_one_in_massapequa, 5),
  updated_replacement_rules
)
```

If the quantity of text being redacted is large, and documents are regularly repeated, the user may wish 
to parse the replacement rules into a caching redaction function:

``` {r}
cached_redacter <- parse_redacter(updated_replacement_rules, with_cache = T)
cached_redacter
```

This new function has a `memoization` layer built in, so that if the same document is presented - replacements are called from memory.  This may speed up data processing if the same passage of text is presented multiple times, but comes at the cost of memory.  The `cached_redact_function` can be used in `redact` in the same way:

``` {r}
redacted_docs <- redact(
  the_one_in_massapequa,
  cached_redacter
)
cached_redacter
```

And its representation tracks the number of unique documents stored.

---

# Summary

The automatic replacement tools allow users to:

- Rapidly generate redaction mappings for large datasets
- Choose between deterministic or randomised replacement strategies
- Convert replacement rules into reusable redaction functions

This approach separates PID identification from transformation, enabling reproducible and auditable redaction workflows.
