---
title: "Automatic Replacement Tools"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Automatic Replacement Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,

  comment = "#>",
  rows.print = 6
)
```


This package intends to aid the user in identifying and removing PID from the data set.
The `report_to_replacement_rules` function is used to generate a `replacement_rules` table, which
contains the rules for replacing PID in the data set.  In the first instance it is intended 
for the user to manually define the `To` column of the `replacement_rules` and then parse the 
file back into R.

If the data set contains a large quantity of PID, and hence it is not practical to
define all replacements, the `auto_replace` utility may be beneficial.  This function
operates on the `replacement_rules` table produced by `report_to_replacement_rules` 
and encodes the `To` column.   The precise replacement method can be defined by the user, 
with three methods supplied by the package:

- `hashing_replacement.f` - hashes the `To` column using a key and salt, 
  producing a unique replacement for each value.
- `random_replacement.f` - replaces the `To` column with a random value from a defined space.
- `all_random_replacement.f` - replaces the `To` column with a random value from a defined space, 
  but ensures that all replacements are unique.
  
The basic workflow is hence:

```{r, results="hide", message=F}
library(pid.pos)

report <- data_frame_report(the_one_in_massapequa)
replacement_rules <- report_to_replacement_rules(report)

replacement_rules
```

``` {r, echo=F}
head(replacement_rules, 5)
```

To fill in the `To` column we initialize a replacement function (in this case `random_replacement.f`):
  
``` {r}
replacement.f <- random_replacement.f(
  replacement_size = 5,
  replacement_space = LETTERS
)
```

and then apply it to the `replacement_rules` table:

``` {r, results="hide"}
set.seed(101)
updated.replacement_rules <- auto_replace(replacement_rules, replacement.f)
updated.replacement_rules
```
``` {r, echo=F}
head(updated.replacement_rules, 5)
```

The `updated.replacement_rules` can then converted to a replacement function via `load_replacement_rules`: 

``` {r, eval=F}
load_replacement_rules(updated.replacement_rules)
```
``` {r, eval=F}
function(.x) dplyr::case_when(
    stringr::str_detect(.x, '(Cut to Monica and Chandler)') ~ .x |> stringr::str_replace_all('Monica', 'CCBTU') |>
      stringr::str_replace_all('Chandler', 'TFXGJ'),
    stringr::str_detect(.x, '(Cut to Monica, at the microphone)') ~ .x |> stringr::str_replace_all('Monica', 'CCBTU'),
    stringr::str_detect(.x, '(Cut to Phoebe and Joey)') ~ .x |> stringr::str_replace_all('Phoebe', 'IYNWQ') |>
      stringr::str_replace_all('Joey', 'JZKUZ'),
    ...,
    .default=.x
  )
```

which if parsed, can be used to modify the original data set:    

``` {r, results="hide", message=F}
redaction.f <- load_replacement_rules(updated.replacement_rules, parse = T)

library(dplyr)
the_one_in_massapequa |>
  mutate(
    across(
      where(is.character),
      redaction.f
    )
  )
```

``` {r, echo = F}
the_one_in_massapequa |>
  mutate(
    across(
      where(is.character),
      redaction.f
    )
  ) |> 
  head(5)
```

