---
title: "Using custom functions in `pid_pos`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using custom functions in `pid_pos`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Out of the box, `pid_pos()` relies on the `udpipe` package to perform part-of-speech (POS) tagging.  
However, the `pid_pos` API is intentionally flexible and allows users to:

- Supply their own custom POS tagging functions
- Implement custom filtering logic
- Integrate alternative tagging engines (e.g., Python libraries)

This vignette demonstrates each of these extensions.

A custom tagger must:

- Accept a single character string (a sentence)
- Return a data frame with one row per token

A custom filter must:

- Accept the tagger’s returned data frame
- Return a filtered version of that data frame

We begin by loading the package:

```{r setup}
library(pid.pos)
```

We will use the example dataset:

```{r}
example.data <- head(the_one_in_massapequa, 20)
example.data
```

---

# Custom POS Tagging Functions

As a minimal example, suppose we want to tag only the main characters from *Friends* as proper nouns.

```{r}
friends <- c("Joey", "Phoebe", "Ross", "Chandler", "Monica", "Rachel")
```

We define a sentence-level tagging function:

```{r}
sentence_tagger <- function(sentence) {

  clean_text <- gsub("[[:punct:]]", "", sentence)
  tokens <- strsplit(clean_text, "\\s+")[[1]]
  tokens <- tokens[tokens != ""]
  
  tibble::tibble(
    Token = tokens,
    upos = ifelse(tokens %in% friends, "PROPN", "XXX")
  )
}

sentence_tagger("Joey sat in Central Perk")
```

To use this with `pid_pos()`, we wrap it with `custom_tagger()`:

```{r}
friends_tagger <- custom_tagger(sentence_tagger)

friends_tagger(example.data$text)
```

We can now supply it to `pid_pos()`:

```{r}
result <- pid_pos(example.data, tagger = friends_tagger)
result
```

---

# Custom Filtering Functions

The default `filter_to_proper_noun()` assumes:

1. Proper nouns are labeled `"PROPN"`
2. The POS column is named `"upos"`

If your tagger uses different conventions, you must supply a custom filter.

For example, suppose:

- Proper nouns are labeled `"NNP"`
- The POS column is named `"POS"`

We define:

```{r}
custom_filter <- function(tag_frm) {
  dplyr::filter(tag_frm, POS == "NNP")
}
```

And a compatible tagger:

```{r}
friends_tagger2 <- custom_tagger(function(sentence) {
  
  if (is.na(sentence)) {
    return(tibble::tibble(Token = character(), POS = character()))
  }
  
  clean_text <- gsub("[[:punct:]]", "", sentence)
  tokens <- strsplit(clean_text, "\\s+")[[1]]
  tokens <- tokens[tokens != ""]
  
  tibble::tibble(
    Token = tokens,
    POS = ifelse(tokens %in% friends, "NNP", "XXX")
  )
})
```

Now use both:

```{r}
pid_pos(
  example.data,
  tagger = friends_tagger2,
  filter = custom_filter
)
```

---

# Using Other Tagging Frameworks (Python Example)

The tagging engine can also be replaced entirely. For example, Python’s `nltk` library provides POS tagging functionality.

The following Python function (included in the package as `nltk_function.py`) defines a simple wrapper:

```python
from nltk.tokenize import word_tokenize
from nltk.tag import pos_tag

def nltk_pos_tagger(sentence):
    tokens = word_tokenize(sentence)
    return pos_tag(tokens)
```

We can bridge this into R using `reticulate` (see [this vignette](https://cran.r-project.org/web/packages/reticulate/vignettes/calling_python.html) for an intro to using `reticulate`).

> **Note:** This section requires a working Python installation and the `nltk` module.  
> If these are not available, the code below will be skipped.

```{r, eval = FALSE}
reticulate::use_virtualenv("YOUR ENVIRONMENT", required = TRUE)

if (!reticulate::py_module_available("nltk")){
  reticulate::py_install("nltk", pip = TRUE)
}
```

```{r, eval = reticulate::py_available(initialize = FALSE)}

nltk_tagger <- custom_tagger(function(sentence) {
  
  tagged <- nltk_pos_tagger(sentence)
  
  tagged_frm <- tibble::as_tibble(
    do.call(rbind, tagged),
    .name_repair = "minimal"
  )
  
  colnames(tagged_frm) <- c("Token", "POS")
  
  dplyr::mutate(
    tagged_frm,
    dplyr::across(dplyr::everything(), as.character)
  )
})
  
nltk_filter <- function(frm) {
  dplyr::filter(frm, POS %in% c("NNP", "NNPS"))
}
  
pid_pos(
  example.data,
  tagger = nltk_tagger,
  filter = nltk_filter
)
```

---

# Summary

The `pid_pos()` API separates:

- **Tagging logic**
- **Filtering logic**

This design allows users to:

- Implement lightweight dictionary-based taggers
- Swap in alternative POS frameworks
- Integrate external engines such as Python libraries
- Customize filtering criteria without modifying tagging code

By adhering to the simple contract:

- Taggers return a token-level data frame
- Filters return a subset of that data frame

`pid_pos()` can operate with virtually any POS tagging system.

