---
title: 'PID.POS: An R package for the detection of personally identifiable data'
bibliography: ../JOSS_utilities/paper.bib
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{PID.POS: An R package for the detection of personally identifiable data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
``` {r setup, echo=F}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
```

# Summary

The `PID.POS` package is designed to aid with the identification of personal identifiability risks in data sets.
By applying existing natural language processing techniques, the package is able to identify proper nouns within a
data set.  The extraction of proper nouns reduced the complexity of the data, allowing for a quicker review and 
oversight of the data.  The package also includes a basic tool for the design, and 
implementation of a redaction process.  

# Statement of need  


The world is embedded in a data revolution.  Never before have we had the depth or breadth of data 
being captured  and analysed than we do at present, and this is only set to increase. 
In response, international bodies are taking steps to ensure legal protection of an individual's rights 
to their own data [@GDPR].  One effect of increase legislation  has been a growing awareness of the role and 
responsibility of data controllers [@ICODataController] and the risks of big data [@clarke2016big]. 
Among these concerns, a risk of 'personal identifiability' i.e. the ability to directly or 
indirectly identify an individual from a dataset [@finck2020they], is paramount  and, if breeched,
can lead to reputation damage and fines [@ICOWhatIf]. 



Where data  is structured  and comprises only a few hundred observations, a manual inspection can 
identify variables which contain directly personally identifiable data (PID) with a reasonable time
investment.  However, in the case of  modern large data sets which may comprise millions of
observations,  a manual inspection may miss PID if it is  embedded within a passage of text, or is
a rarity for the given variable.  The `PID.POS` (Personal Identifiability Detection by Part Of 
Speech tagging) package is designed to aid with the identification of PID risks in data sets.   In
comparison to existing packages which rely on a curated list of common names and string-matching,
`PID.POS` builds on the existing `udpipe` framework, extracting all examples of 
proper nouns and providing a mechanism for the review and redaction of PID risks.



# Comparison to existing R packages

The need to review data sets to identify risks is not new, and there are a number of packages which have been developed
to aid in this process.  The most notable of these are the `PII` package  [@pii], which is designed to 
identify personally identifiable features via pattern matching. These approaches can be effective in 
identifying PID, but have a risk of missing edge cases e.g. relying on sentence case  to identify names. The 
approach taken in `PID.POS` conversely takes the approach of purposefully extracting all proper-nouns, and hence
increase the false positive rate, with the intention of supplying a simplified extract to aid  human interpretation 
rather than fully automate it.

# In practice

To install the current version of `PID.POS` package, use the following code:

``` {r, eval=F}
# install.packages("pak")
pak::pkg_install("Stat-Cook/PID.POS")
```

To assist with understanding the `PID.POS` package, we include a subset of the 'friends' 
data set from the `friends` package.

``` {r, warning = F, message = F, results="hide"}
library(pid.pos)
the_one_in_massapequa
```

``` {r, warning = F, message = F, echo=F}
head(the_one_in_massapequa, 5)
```


The package has two main functions for identifying PID  risks, depending on the users needs.  
First, the `data_frame_report` function converts a typical R data frame into a new data frame of:

* `ID` - the column and row where the sentence first appears
* `Token` - the specific proper noun token 
* `Sentence` - the sentence containing proper nouns
* `Repeats` - the number of times the sentence occurs in the data set
* `Affected Columns` - the columns in the original data frame which contain the sentence

``` {r, results="hide", message=F}
report <- data_frame_report(the_one_in_massapequa)
report
```

``` {r, echo=F, message=F}
head(report, 5)
```

The second function is `report_on_folder` which iterates over a folder of data files,
producing a proper noun report for each.  It is foreseen that this function will 
be the more useful, used just before data release to evidence no PID risks.

``` {r, eval=F}
report_on_folder("path/to/data/")
browse_model_location()
```    

NB: the `data_frame_report` and `report_on_folder` functions automate the download 
of the pre-trained `udpipe` model.  These models are required to be cached to the users 
hard-drive and hence firewall issues may present.  The vignette ... is included to 
help with common issues.

While being able to identify PID risks is the core premise of this package, it would be 
remiss to not supply some tools to aid in the removal of PID.  Hence, we supply  
basic functionality designed for minimal technical knowledge to assist in the 
redaction of PID.

Where a PID report has been ran, the resulting data frame can be passed to the function 
`report_to_replacement_rules` which will convert the report to a csv file with three headings:

* `If` - the sentence pattern which, if it matches, the replacement is applied
* `From` - the pattern to be replaced
* `To` - the intended replacement

``` {r, eval = F, results="hide"}
replacement_rules <- report_to_replacement_rules(
  report,
  path = "path/to/report.csv"
)
```

``` {r, echo=F}
replacement_rules <- report_to_replacement_rules(report)
head(replacement_rules)
```
The csv file is intended to be edited by the data controller, who hence does not need 
to understand R, and can be reimported using the `load_replacement_rules` function:

``` {r, eval=F}
load_replacement_rules("path/to/report.csv")
```

The  `load_replacement_rules` function creates a string replacement rule to capture the 
desired redactions, with the option for R to 'parse' the function for use as part of a data
pipeline:

``` {r, eval=F}
replacement.func <- load_replacement_rules("path/to/report.csv", parse = T)

the_one_in_massapequa |>
  mutate(
    across(
      where(is.character),
      replacement.func
    )
  )
```

For a more advanced user, the `load_replacement_rules` function utilizes the `str_detect` and 
`str_replace_all` functions from the `stringr` package, and hence supports regex. Further utilities are 
available, notably tools to automatically encode the `To` column (see [Auto Replacements](https://stat-cook.github.io/pid.pos/articles/AutoReplacement.html)).

# Current applications

The `PID.POS` package was developed for applications in the NuRS and AmReS research projects
which aim to extract and analyse retrospective operational data from NHS Trusts to understand
staff retention and patient safety.

# Contributions

The package was designed by RC and .... Implementation was done by RC. Quality assurance was 
done by ... Documentation was written by RC.... Funding for the work was won by RC and SJ.

# Acknowledgements

The development of `PID.POS` was part of the NuRS and AmReS projects funded by the
Health Foundation.

# References



