---
title: "Proper Noun Report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{proper_noun_report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `proper_noun_report` function is used to generate a report of proper nouns in a text. The report includes the frequency of each proper noun, the number of times each proper noun is mentioned, and the context in which each proper noun is mentioned.

The function is intended to be used on a vector of sentences, where each sentence is a character string. The function uses the `udpipe` package to identify proper nouns in the text.

## Usage

```{r}
library(pid.pos)

sentence_frm

report <- proper_noun_report(sentence_frm$Sentence)
report$`Proper Nouns`
```

The output of `propoer_noun_report` is a list with three items:

* `Proper Nouns`: a data frame of each "proper noun" detected in the given data.
* `All Tags`: a data frame of every words and tag (the raw output of `udpipe`)
* `Raw Text`: the original text.





